---
title: "gopherDen Database Browser"
subtitle: "Genomic Objects & Provenance for Environmental Research"
date: today
format: html
---

```{r setup}
library(tidyverse)
library(gopheR)
library(knitr)

# Set database path - navigate up from quarto directory
db_path <- normalizePath(file.path("..", "extdata"))
options(gopheR.db_path = db_path)
options(gopheR.db_file = "gopherDen_db.sqlite")
```

## Database Summary

This database contains genomic and environmental data organized using the gopheR framework. Objects are connected through defined relationships (edges) to track provenance and relationships.

```{r object-counts}
# Get object counts by type
object_counts <- gopheR::with_gopher_con(function(con) {
  DBI::dbGetQuery(con, "
    SELECT
      object_type,
      COUNT(*) as count
    FROM object
    GROUP BY object_type
    ORDER BY object_type
  ")
})

kable(object_counts,
      col.names = c("Object Type", "Count"),
      caption = "Objects in Database")
```

```{r edge-counts}
# Get edge counts by type
edge_counts <- gopheR::with_gopher_con(function(con) {
  DBI::dbGetQuery(con, "
    SELECT
      edge_type,
      COUNT(*) as count
    FROM edge
    GROUP BY edge_type
    ORDER BY edge_type
  ")
})

kable(edge_counts,
      col.names = c("Edge Type", "Count"),
      caption = "Relationships in Database")
```

## Study Sites

```{r site-map}
# Get site information
sites <- gopheR::with_gopher_con(function(con) {
  DBI::dbGetQuery(con, "
    SELECT
      object_id as site_id,
      label,
      description
    FROM object
    WHERE object_type = 'site'
  ")
})

# Try to extract lat/long from description
# Format: "Description, 35.9641°N 84.2853°W, elevation..."
sites <- sites |>
  mutate(
    # Extract latitude
    lat = str_extract(description, "\\d+\\.\\d+°N") |>
      str_remove("°N") |>
      as.numeric(),
    # Extract longitude (as negative for W)
    lon = str_extract(description, "\\d+\\.\\d+°W") |>
      str_remove("°W") |>
      as.numeric() |>
      (\(x) -x)()  # Convert West to negative
  )

# Check if we have valid coordinates
if (any(!is.na(sites$lat) & !is.na(sites$lon))) {
  library(leaflet)

  leaflet(sites) |>
    addProviderTiles(providers$CartoDB.Positron) |>
    addCircleMarkers(
      lng = ~lon,
      lat = ~lat,
      radius = 8,
      color = "#2c7fb8",
      fillOpacity = 0.7,
      popup = ~paste0("<strong>", label, "</strong><br>", site_id)
    )
} else {
  cat("No site coordinates available for mapping.")
}
```

## Quick Stats

```{r quick-stats}
# Get some interesting stats
stats <- gopheR::with_gopher_con(function(con) {
  list(
    total_objects = DBI::dbGetQuery(con, "SELECT COUNT(*) as n FROM object")$n,
    total_edges = DBI::dbGetQuery(con, "SELECT COUNT(*) as n FROM edge")$n,
    total_workflows = DBI::dbGetQuery(con, "SELECT COUNT(*) as n FROM workflow")$n,
    total_results = DBI::dbGetQuery(con, "SELECT COUNT(*) as n FROM result")$n
  )
})

tibble(
  Metric = c("Total Objects", "Total Edges", "Total Workflows", "Total Results"),
  Count = c(stats$total_objects, stats$total_edges, stats$total_workflows, stats$total_results)
) |>
  kable()
```

## Explore Further

- [MAGs](mags.html) - Browse metagenomic assembled genomes
- [Samples](samples.html) - View environmental samples
- [Assemblies](assemblies.html) - Explore assembly statistics
