---
title: "Metagenome-Assembled Genomes (MAGs)"
date: today
format: html
---

```{r setup}
library(tidyverse)
library(gopheR)
library(knitr)
library(reactable)

# Set database path
db_path <- normalizePath(file.path("..", "extdata"))
options(gopheR.db_path = db_path)
options(gopheR.db_file = "gopherDen_db.sqlite")
```

## MAG Overview

This page shows all metagenome-assembled genomes (MAGs) in the database with their quality metrics and taxonomy. For detailed historical data, see example detail pages:

- [MAG_A1_001 Detail Page](objects/mag-MAG_A1_001.html) - Shows all historical quality assessments, taxonomy updates, provenance, and files

```{r mag-data}
# Get all MAGs with their quality metrics
mags <- gopheR::with_gopher_con(function(con) {
  DBI::dbGetQuery(con, "
    SELECT
      o.object_id,
      o.label,
      o.description,
      o.created_at
    FROM object o
    WHERE o.object_type = 'genome'
      AND o.object_subtype = 'MAG'
    ORDER BY o.object_id
  ")
})

# Get MOST RECENT result data for MAGs (join with workflow to get dates)
results <- gopheR::with_gopher_con(function(con) {
  DBI::dbGetQuery(con, "
    SELECT
      r.object_id,
      r.key,
      r.value,
      r.workflow_id,
      w.workflow_date
    FROM result r
    JOIN workflow w ON w.workflow_id = r.workflow_id
    WHERE r.object_id LIKE 'MAG%'
    ORDER BY r.object_id, r.key, w.workflow_date DESC
  ")
})

# For landing page, keep only the MOST RECENT value per object_id + key
results_latest <- results |>
  group_by(object_id, key) |>
  slice_max(workflow_date, n = 1, with_ties = FALSE) |>
  ungroup()

# Pivot results wide
results_wide <- results_latest |>
  select(object_id, key, value) |>
  pivot_wider(
    names_from = key,
    values_from = value
  )

# Join with MAG metadata
mags_full <- mags |>
  left_join(results_wide, by = "object_id")
```

## MAG Quality Summary

```{r mag-table}
# Only convert numeric columns that exist
mags_display <- mags_full
if ("completeness" %in% colnames(mags_display)) {
  mags_display <- mags_display |> mutate(completeness = as.numeric(completeness))
}
if ("contamination" %in% colnames(mags_display)) {
  mags_display <- mags_display |> mutate(contamination = as.numeric(contamination))
}
if ("size_bp" %in% colnames(mags_display)) {
  mags_display <- mags_display |> mutate(size_bp = as.numeric(size_bp))
}
if ("gc_content" %in% colnames(mags_display)) {
  mags_display <- mags_display |> mutate(gc_content = as.numeric(gc_content))
}
if ("n_contigs" %in% colnames(mags_display)) {
  mags_display <- mags_display |> mutate(n_contigs = as.numeric(n_contigs))
}
if ("N50" %in% colnames(mags_display)) {
  mags_display <- mags_display |> mutate(N50 = as.numeric(N50))
}

# Select only columns that exist
select_cols <- c("object_id", "label")
if ("completeness" %in% colnames(mags_display)) select_cols <- c(select_cols, "completeness")
if ("contamination" %in% colnames(mags_display)) select_cols <- c(select_cols, "contamination")
if ("size_bp" %in% colnames(mags_display)) select_cols <- c(select_cols, "size_bp")
if ("n_contigs" %in% colnames(mags_display)) select_cols <- c(select_cols, "n_contigs")
if ("N50" %in% colnames(mags_display)) select_cols <- c(select_cols, "N50")
if ("gc_content" %in% colnames(mags_display)) select_cols <- c(select_cols, "gc_content")
if ("taxonomy" %in% colnames(mags_display)) select_cols <- c(select_cols, "taxonomy")

mags_display <- mags_display |> select(all_of(select_cols))

# Build column definitions dynamically
col_defs <- list(
  object_id = colDef(name = "MAG ID", minWidth = 120),
  label = colDef(name = "Label", minWidth = 150)
)

if ("completeness" %in% colnames(mags_display)) {
  col_defs$completeness <- colDef(
    name = "Completeness (%)",
    format = colFormat(digits = 1),
    style = function(value) {
      if (!is.na(value) && value >= 90) {
        list(color = "#008000", fontWeight = "bold")
      } else if (!is.na(value) && value >= 50) {
        list(color = "#ff8c00")
      } else {
        list(color = "#999")
      }
    }
  )
}

if ("contamination" %in% colnames(mags_display)) {
  col_defs$contamination <- colDef(
    name = "Contamination (%)",
    format = colFormat(digits = 1),
    style = function(value) {
      if (!is.na(value) && value <= 5) {
        list(color = "#008000")
      } else if (!is.na(value) && value <= 10) {
        list(color = "#ff8c00")
      } else {
        list(color = "#d32f2f", fontWeight = "bold")
      }
    }
  )
}

if ("size_bp" %in% colnames(mags_display)) {
  col_defs$size_bp <- colDef(
    name = "Size (bp)",
    format = colFormat(separators = TRUE, digits = 0)
  )
}

if ("n_contigs" %in% colnames(mags_display)) {
  col_defs$n_contigs <- colDef(name = "Contigs", format = colFormat(digits = 0))
}

if ("N50" %in% colnames(mags_display)) {
  col_defs$N50 <- colDef(name = "N50", format = colFormat(separators = TRUE, digits = 0))
}

if ("gc_content" %in% colnames(mags_display)) {
  col_defs$gc_content <- colDef(name = "GC %", format = colFormat(digits = 1))
}

if ("taxonomy" %in% colnames(mags_display)) {
  col_defs$taxonomy <- colDef(name = "Taxonomy", minWidth = 200)
}

reactable(
  mags_display,
  searchable = TRUE,
  filterable = TRUE,
  striped = TRUE,
  highlight = TRUE,
  defaultPageSize = 20,
  columns = col_defs
)
```

## Quality Distribution

```{r quality-plot}
mags_full |>
  mutate(
    completeness = as.numeric(completeness),
    contamination = as.numeric(contamination)
  ) |>
  filter(!is.na(completeness), !is.na(contamination)) |>
  ggplot(aes(x = completeness, y = contamination)) +
  geom_point(size = 3, alpha = 0.7, color = "#2c7fb8") +
  geom_hline(yintercept = 5, linetype = "dashed", color = "darkgreen", alpha = 0.5) +
  geom_hline(yintercept = 10, linetype = "dashed", color = "orange", alpha = 0.5) +
  geom_vline(xintercept = 90, linetype = "dashed", color = "darkgreen", alpha = 0.5) +
  geom_vline(xintercept = 50, linetype = "dashed", color = "orange", alpha = 0.5) +
  annotate("text", x = 95, y = 2, label = "High-quality", color = "darkgreen", hjust = 0) +
  annotate("text", x = 70, y = 7, label = "Medium-quality", color = "orange") +
  labs(
    title = "MAG Quality Assessment",
    subtitle = "Based on completeness and contamination",
    x = "Completeness (%)",
    y = "Contamination (%)"
  ) +
  theme_minimal() +
  theme(
    plot.title = element_text(size = 14, face = "bold"),
    plot.subtitle = element_text(size = 11, color = "gray40")
  )
```

## Size Distribution

```{r size-histogram}
mags_full |>
  mutate(size_mb = as.numeric(size_bp) / 1e6) |>
  filter(!is.na(size_mb)) |>
  ggplot(aes(x = size_mb)) +
  geom_histogram(bins = 20, fill = "#2c7fb8", alpha = 0.7, color = "white") +
  labs(
    title = "MAG Genome Size Distribution",
    x = "Genome Size (Mb)",
    y = "Count"
  ) +
  theme_minimal() +
  theme(plot.title = element_text(size = 14, face = "bold"))
```

## Taxonomy Overview

```{r taxonomy-summary}
mags_full |>
  filter(!is.na(taxonomy)) |>
  separate(taxonomy, into = c("domain", "phylum", "class", "order", "family", "genus", "species"),
           sep = ";", fill = "right", extra = "drop") |>
  mutate(phylum = str_trim(phylum)) |>
  count(phylum, sort = TRUE) |>
  filter(!is.na(phylum)) |>
  ggplot(aes(x = reorder(phylum, n), y = n)) +
  geom_col(fill = "#2c7fb8", alpha = 0.8) +
  coord_flip() +
  labs(
    title = "MAGs by Phylum",
    x = NULL,
    y = "Count"
  ) +
  theme_minimal() +
  theme(plot.title = element_text(size = 14, face = "bold"))
```

## Summary Statistics

```{r summary-stats}
summary_stats <- mags_full |>
  summarise(
    `Total MAGs` = n(),
    `High-quality (>90% complete, <5% contam)` = sum(
      as.numeric(completeness) > 90 & as.numeric(contamination) < 5, na.rm = TRUE
    ),
    `Medium-quality (>50% complete, <10% contam)` = sum(
      as.numeric(completeness) > 50 & as.numeric(completeness) <= 90 &
      as.numeric(contamination) < 10, na.rm = TRUE
    ),
    `Mean Completeness (%)` = round(mean(as.numeric(completeness), na.rm = TRUE), 1),
    `Mean Contamination (%)` = round(mean(as.numeric(contamination), na.rm = TRUE), 1),
    `Mean Size (Mb)` = round(mean(as.numeric(size_bp), na.rm = TRUE) / 1e6, 2)
  )

summary_stats |>
  pivot_longer(everything(), names_to = "Metric", values_to = "Value") |>
  kable()
```
