---
title: "MAG Detail: MAG_A1_001"
date: today
format: html
params:
  mag_id: "MAG_A1_001"
---

```{r setup}
library(tidyverse)
library(gopheR)
library(knitr)
library(reactable)

# Set database path
db_path <- normalizePath(file.path("..", "..", "extdata"))
options(gopheR.db_path = db_path)
options(gopheR.db_file = "gopherDen_db.sqlite")

mag_id <- params$mag_id
```

## MAG Metadata

```{r mag-metadata}
mag <- gopheR::with_gopher_con(function(con) {
  DBI::dbGetQuery(con, "
    SELECT
      object_id,
      object_type,
      object_subtype,
      label,
      description,
      created_at,
      created_by
    FROM object
    WHERE object_id = ?
  ", params = list(mag_id))
})

kable(
  mag |> pivot_longer(everything(), names_to = "Field", values_to = "Value"),
  col.names = c("Field", "Value")
)
```

## Quality Metrics - Historical

This table shows ALL quality assessments for this MAG over time. The most recent assessment is at the top.

```{r quality-history}
quality_results <- gopheR::with_gopher_con(function(con) {
  DBI::dbGetQuery(con, "
    SELECT
      r.key,
      r.value,
      r.workflow_id,
      w.description as workflow_description,
      w.workflow_date,
      w.created_by
    FROM result r
    JOIN workflow w ON w.workflow_id = r.workflow_id
    WHERE r.object_id = ?
      AND r.key IN ('completeness', 'contamination', 'size_bp', 'gc_content', 'N50', 'n_contigs')
    ORDER BY r.key, w.workflow_date DESC
  ", params = list(mag_id))
})

if (nrow(quality_results) > 0) {
  reactable(
    quality_results,
    groupBy = "key",
    searchable = TRUE,
    striped = TRUE,
    defaultExpanded = TRUE,
    columns = list(
      key = colDef(name = "Metric", minWidth = 120),
      value = colDef(name = "Value", minWidth = 100),
      workflow_id = colDef(name = "Workflow ID", minWidth = 120),
      workflow_date = colDef(name = "Date", minWidth = 120),
      workflow_description = colDef(name = "Description", minWidth = 200),
      created_by = colDef(name = "Analyst", minWidth = 100)
    )
  )
} else {
  cat("No quality metrics available for this MAG.")
}
```

## Taxonomy - Historical

This table shows ALL taxonomic classifications for this MAG over time. Taxonomy may change as reference databases are updated (e.g., GTDB releases).

```{r taxonomy-history}
taxonomy_results <- gopheR::with_gopher_con(function(con) {
  DBI::dbGetQuery(con, "
    SELECT
      r.value as taxonomy,
      r.workflow_id,
      w.description as workflow_description,
      w.workflow_date,
      w.created_by
    FROM result r
    JOIN workflow w ON w.workflow_id = r.workflow_id
    WHERE r.object_id = ?
      AND r.key = 'taxonomy'
    ORDER BY w.workflow_date DESC
  ", params = list(mag_id))
})

if (nrow(taxonomy_results) > 0) {
  reactable(
    taxonomy_results,
    searchable = TRUE,
    striped = TRUE,
    highlight = TRUE,
    columns = list(
      taxonomy = colDef(name = "Taxonomy", minWidth = 300),
      workflow_id = colDef(name = "Workflow ID", minWidth = 120),
      workflow_date = colDef(name = "Date", minWidth = 120),
      workflow_description = colDef(name = "Description", minWidth = 200),
      created_by = colDef(name = "Analyst", minWidth = 100)
    )
  )
} else {
  cat("No taxonomy data available for this MAG.")
}
```

## Current Best Values

For quick reference, here are the MOST RECENT values:

```{r current-values}
current_results <- gopheR::with_gopher_con(function(con) {
  DBI::dbGetQuery(con, "
    SELECT
      r.key,
      r.value,
      w.workflow_date,
      r.workflow_id
    FROM result r
    JOIN workflow w ON w.workflow_id = r.workflow_id
    WHERE r.object_id = ?
    ORDER BY r.key, w.workflow_date DESC
  ", params = list(mag_id))
})

current_best <- current_results |>
  group_by(key) |>
  slice_max(workflow_date, n = 1, with_ties = FALSE) |>
  ungroup() |>
  select(key, value, workflow_date, workflow_id)

kable(current_best,
      col.names = c("Metric", "Value", "As of", "Workflow ID"))
```

## Provenance: Assembly & Readsets

Shows how this MAG was created.

```{r provenance}
# Get binned_from relationship (MAG <- assembly)
binned_from <- gopheR::with_gopher_con(function(con) {
  DBI::dbGetQuery(con, "
    SELECT
      e.parent_id as assembly_id,
      o.label as assembly_label
    FROM edge e
    JOIN object o ON o.object_id = e.parent_id
    WHERE e.child_id = ?
      AND e.edge_type = 'binned_from'
  ", params = list(mag_id))
})

if (nrow(binned_from) > 0) {
  cat("### Binned From Assembly\n\n")
  kable(binned_from, col.names = c("Assembly ID", "Label"))

  assembly_id <- binned_from$assembly_id[1]

  # Get readsets used for assembly (assembled_from)
  assembled_from <- gopheR::with_gopher_con(function(con) {
    DBI::dbGetQuery(con, "
      SELECT
        e.parent_id as readset_id,
        o.label as readset_label
      FROM edge e
      JOIN object o ON o.object_id = e.parent_id
      WHERE e.child_id = ?
        AND e.edge_type = 'assembled_from'
    ", params = list(assembly_id))
  })

  if (nrow(assembled_from) > 0) {
    cat("\n### Assembly Built From Readsets\n\n")
    kable(assembled_from, col.names = c("Readset ID", "Label"))
  }
}

# Get binned_with relationship (MAG <- readsets for coverage)
binned_with <- gopheR::with_gopher_con(function(con) {
  DBI::dbGetQuery(con, "
    SELECT
      e.parent_id as readset_id,
      o.label as readset_label
    FROM edge e
    JOIN object o ON o.object_id = e.parent_id
    WHERE e.child_id = ?
      AND e.edge_type = 'binned_with'
  ", params = list(mag_id))
})

if (nrow(binned_with) > 0) {
  cat("\n### Binned Using Coverage From Readsets\n\n")
  kable(binned_with, col.names = c("Readset ID", "Label"))
}
```

## Observations: Where This MAG Was Found

Shows which samples this MAG has been observed in (post-hoc detection).

```{r observations}
observations <- gopheR::with_gopher_con(function(con) {
  DBI::dbGetQuery(con, "
    SELECT
      e.parent_id as readset_id,
      rs.label as readset_label
    FROM edge e
    JOIN object rs ON rs.object_id = e.parent_id
    WHERE e.child_id = ?
      AND e.edge_type = 'observed_in'
  ", params = list(mag_id))
})

if (nrow(observations) > 0) {
  kable(observations, col.names = c("Readset ID", "Label"))
} else {
  cat("No observation data available (MAG not mapped to other samples).")
}
```

## Files

All files associated with this MAG.

```{r files}
files <- gopheR::with_gopher_con(function(con) {
  DBI::dbGetQuery(con, "
    SELECT
      file_role,
      file_path,
      file_format,
      checksum,
      workflow_id
    FROM object_file
    WHERE object_id = ?
    ORDER BY file_role
  ", params = list(mag_id))
})

if (nrow(files) > 0) {
  kable(files)
} else {
  cat("No files registered for this MAG.")
}
```
